#Course 5 Reproducesable Research - Week 2 Assignment 
###by Hari Sriraman

#Peer-graded Assignment: Course Project 1

### Code for reading in the dataset and/or processing the data

```{r Read File, echo=TRUE}
setwd("C:/Users/hsrirama/Documents/Personal/Coursera/Course 5 - Reproducesable Research/Project")

if(!file.exists("activity.csv")) {
    unzip("repdata_data_activity.zip")
}

acty_ds = read.csv("activity.csv")

```

### Histogram of the total number of steps taken each day
```{r Histogram ~ Date, echo=TRUE}
acty_steps_by_day <- aggregate(steps ~ date, acty_ds, sum, na.action = na.pass)
hist(acty_steps_by_day$steps, col="blue", main = "Total # steps/day", xlab = "# of steps")

```

### Mean and median number of steps taken each day
```{r Mean and Median steps/day}
r_stp_mean <- mean(acty_steps_by_day$steps)
message("Mean = ", r_stp_mean)

r_stp_medn <- median(acty_steps_by_day$steps)
message("Median = ", r_stp_medn)
```

### Time series plot of the average number of steps taken. 

```{r time series}
acty_mean_interval <- tapply(acty_ds$steps, acty_ds$interval, mean, na.rm = T)

plot(acty_mean_interval, type = "l", main = "Time Series Plot", xlab = "The 5 minute interval", ylab = "the Avg # of steps")

```

### The 5-minute interval that, on average, contains the maximum number of steps

```{r max # of steps}
head(sort(acty_mean_interval, decreasing = TRUE), 1)
```

### Code to describe and show a strategy for imputing missing data

```{r}
acty_stepsNA <- sum(is.na(acty_ds$steps))
acty_dateNA <- sum(is.na(acty_ds$date))
acty_intvlNA <- sum(is.na(acty_ds$interval))

message("Missing Steps = ", acty_stepsNA)
message("Missing Dates = ", acty_dateNA)
message("Missing Interval = ", acty_intvlNA)

table(is.na(acty_ds$steps), acty_ds$interval, acty_ds$date)
```

### Histogram of the total number of steps taken each day after missing values are imputed
```{r}
acty_steps_by_day <- aggregate(steps ~ date, acty_ds, sum, na.action = na.exclude)
hist(acty_steps_by_day$steps, col="blue", main = "Total # steps/day", xlab = "# of steps")
```

### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r comparison}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
acty_ds$dow = as.factor(ifelse(is.element(weekdays(as.Date(acty_ds$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_acty <- aggregate(steps ~ interval + dow, acty_ds, mean)

library(lattice)

xyplot(steps_by_interval_acty$steps ~ steps_by_interval_acty$interval|steps_by_interval_acty$dow, main="Average steps/day by interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```

### End of document
